\section{Tomek link complexity measure}
Here we define a measure based on the Tomek Links \JG{Citation} commonly used in undersampling\JG{Citation}. Tomek Links are pairs of mutual nearest neighbors of opposite classes, as described in Equation \ref{eq:tomek_links} where $x_i$ and $x_j$ are instances belonging to class $y_i$ and $y_j$. The point $x_k$ is an arbitrary point in the dataset. $C_{min}$ is set of minority class instances. 
\begin{align}
    d(x_i,x_j) <&\ d(x_i,x_k) \label{eq:tomek_links} \\
    \nonumber d(x_i,x_j) <&\ d(x_j,x_k) \\
    \nonumber y_i=C_{min} \lor&\ y_j = C_{min} \\
    \nonumber y_i \neq&\ y_j 
\end{align}
Given the definition of Tomek Links in Equation \ref{eq:tomek_links} the measure proposed counts the number of Tomek Links, it has to remove, or can remove in order to achieve a uniform class distribution. It is important to notice, that the number of Tomek Links one can remove to achieve a uniform class distribution can be anything from $0$ if there is no Tomek Links in the dataset and up to the number of points, which makes the distribution uniform.

\begin{equation}
    \label{eq:number_of_tomek_links}
    \text{TLS} = \sum_{i \in C_{min}}\ \sum_{j \in \mathcal{D} \setminus C_{min}} TL(x_i,x_j)
\end{equation}

\begin{equation}
    \label{eq:tlcm}
    \text{TLCM}(x)= 
\begin{cases}
    \frac{\text{TLS}}{|\mathcal{D}|},\text{if } |C_{maj}|-|C_{min}| > \text{TLS}\\
    |C_{maj}|-|C_{min}|, \text{otherwise}
\end{cases}
\end{equation}
In Equation \ref{eq:number_of_tomek_links} the number of Tomek links in the dataset is defined as $\text{TLS}$. Here $\mathcal{D}$ is the set of all instances in the domain and $TL$ is the identity function that returns $1$, if $x_i$ and $x_j$ form a Tomek link according to definition \ref{eq:tomek_links}. The cardinality $|C_{min}|$ defines the number of points within the set of minority instances. The measure proposed in this paper is defined as the Tomek Link Complexity Measure ($\text{TLCM}$) and is defined in Equation \ref{eq:tlcm}.
Tomek Links makes the class boundary more complex, so $\text{TLCM}$ reports the percentage of Tomek Links it is possible to remove. The measure might be negatively correlated with the imbalance ratio, when the ratio gets extreme. 
\JG{Insert a class distribution histogram, and gives us an idea about the measure. How big a percentage of the majority class can be removed.}



